Left-linear rules in adhesive categories

### 2024-02-14 icalp/fusione.tex
minima

The previous definition might sound restrictive. However, it is immediate to see that linear rewriting systems are all well-switching (see Proposition~\ref{pr:weak}).
%
Moreover, Lemma~\ref{bono} will show that a class of rewriting systems, comprising
all the ones used in concrete modelling applications of DPO rewriting
(e.g., in the encoding of process calculi, modelling of bio and chemical systems) are actually well-switching.\todo{citazioni?}
%
Roughly speakings, these are rewriting systems over possibly
hierarchical graphical structures obtained as algebras of suitable
signatures where rules are constrained not to merge elements of top
level sorts in the hierarchy (for graphs, nodes can be merged while
edges can not). We first introduce the signature which is essentially
an acyclic graph where nodes can be seen as sorts.

@@

\begin{definition}[Node-merging graphical rewriting systems]
  Let $\mathcal{G}=(V, E, s, t)$ be a finite directed acyclic graph, the
  category $\gph{G}$ of \emph{$\mathcal{G}$-graphs} is the category
  $\Set^{\G}$, where $\G$ is the free category over $\mathcal{G}$. A
  \emph{node-merging graphical rewriting systems} is a left-linear
  rewriting system $(\gph{G}, \R)$ such that, for every rule
  $(l,r)$, with $l\colon K\to L$ and $r\colon K\to R$ in $\R$, the
  following hold
  \begin{enumerate}
  \item for every $v\in V$ and $x\in L(v)$, there exists a vertex $w$
    with depth $0$ and an arrow $p\colon w\to v$ (i.e.~a path in
    $\mathcal{G}$) such that $x$ belongs to the image of
    $L(p)\colon L(w)\to L(v)$;
  \item $r\colon K\to R$ is mono on vertices of depth $0$, i.e. for every
    $v\in V$ such that $\dph(v)=0$, the component $r_v:K(v)\to R(v)$
    is mono.
  \end{enumerate}
\end{definition}

>>>>

The previous definition might sound restrictive. However, it is immediate to see that linear rewriting systems are all well-switching (see Proposition~\ref{pr:weak}).
%
Moreover, Lemma~\ref{bono} will show that a class of rewriting systems, comprising
all the ones used in modelling the applications to the encoding of process calculi or of bio and chemical systems mentioned in the Introduction,
are actually well-switching.
%\todo{citazioni?}
%
Roughly speakings, these are rewriting systems over possibly
hierarchical graphical structures obtained as algebras of suitable
signatures where rules are constrained not to merge elements of top
level sorts in the hierarchy (for graphs, nodes can be merged while
edges can not). We first introduce the signature which is essentially
an acyclic graph where nodes can be seen as sorts.

@@

\begin{definition}[Node-merging graphical rewriting systems]
  Let $\mathcal{G}=(V, E, s, t)$ be a finite directed acyclic graph
  and $\G$ the free category over $\mathcal{G}$. The
  category $\gph{G}$ of \emph{$\mathcal{G}$-graphs} is the category
  $\Set^{\G}$. A
  \emph{node-merging graphical rewriting systems} is a left-linear
  rewriting system $(\gph{G}, \R)$ such that, for every rule
  $(l,r)$, with $l\colon K\to L$ and $r\colon K\to R$ in $\R$, the
  following hold
  \begin{enumerate}
  \item for every $v\in V$ and $x\in L(v)$, there exists a vertex $w$
    with depth $0$ and an arrow $p\colon w\to v$ (i.e.~a path in
    $\mathcal{G}$) such that $x$ belongs to the image of
    $L(p)\colon L(w)\to L(v)$;
  \item $r\colon K\to R$ is mono on vertices of depth $0$, i.e. for every
    $v\in V$ such that $\dph(v)=0$, the component $r_v:K(v)\to R(v)$
    is mono.
  \end{enumerate}
\end{definition}
